### 2023-12-22 assets/simulation/ic_rate_user.tex
:zap: Refactor code

% This file was created by matlab2tikz.
%
%The latest updates can be retrieved from
%  http://www.mathworks.com/matlabcentral/fileexchange/22022-matlab2tikz-matlab2tikz
%where you can also make suggestions and rate matlab2tikz.
%
\definecolor{mycolor1}{rgb}{0.00000,0.44706,0.74118}%
\definecolor{mycolor2}{rgb}{0.85098,0.32549,0.09804}%
%
\begin{tikzpicture}

\begin{axis}[%
width=9.509cm,
height=7.5cm,
at={(0cm,0cm)},
scale only axis,
xmin=2,
xmax=10,
xlabel style={font=\color{white!15!black}},
xlabel={User Pairs},
ymin=24.0366620568765,
ymax=62.9320178950178,
ylabel style={font=\color{white!15!black}},
ylabel={Weighted Sum-Rate [bit/s/Hz]},
axis background/.style={fill=white},
xmajorgrids,
ymajorgrids,
legend style={at={(0.03,0.97)}, anchor=north west, legend cell align=left, align=left, draw=white!15!black}
]
\addplot [color=black, line width=2.0pt]
  table[row sep=crcr]{%
2	24.0366620568765\\
4	34.2503871208668\\
6	41.622749643327\\
8	44.0380050877313\\
10	43.8795026226494\\
};
\addlegendentry{$N^\mathrm{S} = 0$}

\addplot [color=mycolor1, line width=2.0pt, mark=o, mark options={solid, mycolor1}]
  table[row sep=crcr]{%
2	25.3990077780911\\
4	36.3182517013294\\
6	42.844297578367\\
8	46.9811600018883\\
10	47.4095039520004\\
};
\addlegendentry{$(N^\mathrm{S}, L) = (16, 1)$}

\addplot [color=mycolor1, dashed, line width=2.0pt, mark=o, mark options={solid, mycolor1}]
  table[row sep=crcr]{%
2	26.1302675490509\\
4	37.2433314642391\\
6	43.8241685215847\\
8	48.7515778844427\\
10	50.0262345258606\\
};
\addlegendentry{$(N^\mathrm{S}, L) = (16, 4)$}

\addplot [color=mycolor1, dotted, line width=2.0pt, mark=o, mark options={solid, mycolor1}]
  table[row sep=crcr]{%
2	26.5005433882963\\
4	37.8529809047498\\
6	44.5290585688344\\
8	50.1996330401944\\
10	51.8865599593385\\
};
\addlegendentry{$(N^\mathrm{S}, L) = (16, 16)$}

\addplot [color=mycolor2, line width=2.0pt, mark=+, mark options={solid, mycolor2}]
  table[row sep=crcr]{%
2	29.0139251001163\\
4	39.8325157791058\\
6	47.087891457604\\
8	52.0445486861005\\
10	53.9856612942705\\
};
\addlegendentry{$(N^\mathrm{S}, L) = (64, 1)$}

\addplot [color=mycolor2, dashed, line width=2.0pt, mark=+, mark options={solid, mycolor2}]
  table[row sep=crcr]{%
2	31.3276950788263\\
4	42.7749665094192\\
6	50.0160405682383\\
8	56.5092293329805\\
10	59.5047215843355\\
};
\addlegendentry{$(N^\mathrm{S}, L) = (64, 4)$}

\addplot [color=mycolor2, dotted, line width=2.0pt, mark=+, mark options={solid, mycolor2}]
  table[row sep=crcr]{%
2	32.6664816559571\\
4	44.5042904532228\\
6	51.6271886078106\\
8	58.8237458788222\\
10	62.9320178950178\\
};
\addlegendentry{$(N^\mathrm{S}, L) = (64, 64)$}

\end{axis}
\end{tikzpicture}%