### 2023-12-22 assets/simulation/pc_singular_bound.tex

% This file was created by matlab2tikz.
%
%The latest updates can be retrieved from
%  http://www.mathworks.com/matlabcentral/fileexchange/22022-matlab2tikz-matlab2tikz
%where you can also make suggestions and rate matlab2tikz.
%
\definecolor{mycolor1}{rgb}{0.58039,0.76863,0.96078}%
\definecolor{mycolor2}{rgb}{0.25882,0.37647,0.66667}%
\definecolor{mycolor3}{rgb}{0.94902,0.38039,0.38039}%
\definecolor{mycolor4}{rgb}{0.91765,0.03529,0.03529}%
\definecolor{mycolor5}{rgb}{0.46667,0.67451,0.18824}%
%
\begin{tikzpicture}

\begin{axis}[%
width=9.509cm,
height=7.5cm,
at={(0cm,0cm)},
scale only axis,
bar shift auto,
xmin=0,
xmax=5,
xtick={1,2,3,4},
xticklabels={{$\sigma_1$},{$\sigma_2$},{$\sigma_3$},{$\sigma_4$}},
ymin=0,
ymax=4.5,
ylabel style={font=\color{white!15!black}},
ylabel={Amplitude},
axis background/.style={fill=white},
xmajorgrids,
ymajorgrids,
legend style={legend cell align=left, align=left, draw=white!15!black},
legend columns=5,
transpose legend,
legend style={/tikz/column 2/.style={column sep=5pt}}
]
\addplot[ybar, bar width=0.123, fill=black, draw=black, area legend] table[row sep=crcr] {%
1	3.11578483462081\\
2	2.33621539044729\\
3	1.33501504737837\\
4	0.346232804852115\\
};
\addplot[forget plot, color=white!15!black] table[row sep=crcr] {%
0	0\\
5	0\\
};
\addlegendentry{Direct}

\addplot[ybar, bar width=0.123, fill=mycolor1, draw=black, area legend] table[row sep=crcr] {%
1	2.74905207049567\\
2	1.83843449000086\\
3	0.952844790933556\\
4	0.022527302733258\\
};
\addplot[forget plot, color=white!15!black] table[row sep=crcr] {%
0	0\\
5	0\\
};
\addlegendentry{Min, $L = 1$}

\addplot[ybar, bar width=0.123, fill=mycolor2, draw=black, area legend] table[row sep=crcr] {%
1	2.71936092735902\\
2	1.81605826060481\\
3	0.906339746943746\\
4	0.00914697809349159\\
};
\addplot[forget plot, color=white!15!black] table[row sep=crcr] {%
0	0\\
5	0\\
};
\addlegendentry{Min, $L = 32$}

\addplot[ybar, bar width=0.123, fill=mycolor3, draw=black, area legend] table[row sep=crcr] {%
1	3.8960081771894\\
2	2.67136418131802\\
3	1.64221742934803\\
4	0.603919247269453\\
};
\addplot[forget plot, color=white!15!black] table[row sep=crcr] {%
0	0\\
5	0\\
};
\addlegendentry{Max, $L = 1$}

\addplot[ybar, bar width=0.123, fill=mycolor4, draw=black, area legend] table[row sep=crcr] {%
1	4.03127940593837\\
2	2.67500105747776\\
3	1.67178857361767\\
4	0.628233917896107\\
};
\addplot[forget plot, color=white!15!black] table[row sep=crcr] {%
0	0\\
5	0\\
};
\addlegendentry{Max, $L = 32$}

\addplot [color=mycolor5, line width=2.0pt]
  table[row sep=crcr]{%
0	2.68074858002325\\
5	2.68074858002325\\
};
\addlegendentry{$\sigma_1(\mathbf{T})$}

\addplot [color=mycolor5, dashed, line width=2.0pt]
  table[row sep=crcr]{%
0	1.7029380545254\\
5	1.7029380545254\\
};
\addlegendentry{$\sigma_2(\mathbf{T})$}

\addplot [color=mycolor5, dotted, line width=2.0pt]
  table[row sep=crcr]{%
0	0.745179150575099\\
5	0.745179150575099\\
};
\addlegendentry{$\sigma_3(\mathbf{T})$}

\addplot [color=mycolor5, dashdotted, line width=2.0pt]
  table[row sep=crcr]{%
0	1.47492426026826e-08\\
5	1.47492426026826e-08\\
};
\addlegendentry{$\sigma_4(\mathbf{T})$}

\end{axis}

\begin{axis}[%
width=12.27cm,
height=9.202cm,
at={(-1.595cm,-1.012cm)},
scale only axis,
xmin=0,
xmax=1,
ymin=0,
ymax=1,
axis line style={draw=none},
ticks=none,
axis x line*=bottom,
axis y line*=left,
legend columns=5,
transpose legend,
legend style={/tikz/column 2/.style={column sep=5pt}}
]
\end{axis}
\end{tikzpicture}%