### 2024-02-14 assets/simulation/power_sx_txrx16.tex
:sparkles: Refactor code

% This file was created by matlab2tikz.
%
%The latest updates can be retrieved from
%  http://www.mathworks.com/matlabcentral/fileexchange/22022-matlab2tikz-matlab2tikz
%where you can also make suggestions and rate matlab2tikz.
%
\definecolor{mycolor1}{rgb}{0.00000,0.44706,0.74118}%
\definecolor{mycolor2}{rgb}{0.85098,0.32549,0.09804}%
\definecolor{mycolor3}{rgb}{0.92941,0.69412,0.12549}%
\definecolor{mycolor4}{rgb}{0.49412,0.18431,0.55686}%
\definecolor{mycolor5}{rgb}{0.46667,0.67451,0.18824}%
%
\begin{tikzpicture}

\begin{axis}[%
width=7.607cm,
height=6cm,
at={(0cm,0cm)},
scale only axis,
xmin=1,
xmax=9,
xtick={1,2,3,4,5,6,7,8,9},
xticklabels={{$2^0$},{$2^1$},{$2^2$},{$2^3$},{$2^4$},{$2^5$},{$2^6$},{$2^7$},{$2^8$}},
xlabel style={font=\color{white!15!black}},
xlabel={RIS Group Size},
ymode=log,
ymin=7.63413889612284e-05,
ymax=0.000357009156162705,
yminorticks=true,
ylabel style={font=\color{white!15!black}},
ylabel={Channel Power [W]},
axis background/.style={fill=white},
xmajorgrids,
ymajorgrids,
yminorgrids,
legend style={at={(0.03,0.97)}, anchor=north west, legend cell align=left, align=left, draw=white!15!black}
]
\addplot [color=black, line width=2.0pt]
  table[row sep=crcr]{%
1	8.03593568012931e-05\\
2	8.03593568012931e-05\\
3	8.03593568012931e-05\\
4	8.03593568012931e-05\\
5	8.03593568012931e-05\\
6	8.03593568012931e-05\\
7	8.03593568012931e-05\\
8	8.03593568012931e-05\\
9	8.03593568012931e-05\\
};
\addlegendentry{No RIS}

\addplot [color=mycolor1, line width=2.0pt, mark=o, mark options={solid, mycolor1}]
  table[row sep=crcr]{%
1	8.32187269864919e-05\\
2	8.41922087195748e-05\\
3	8.55795092915898e-05\\
4	8.71382957344062e-05\\
5	8.87320958720005e-05\\
};
\addlegendentry{$N_\mathrm{S} = 16$}

\addplot [color=mycolor2, dashed, line width=2.0pt, mark=+, mark options={solid, mycolor2}]
  table[row sep=crcr]{%
1	8.6466574171426e-05\\
2	8.81326989213474e-05\\
3	9.06190783177407e-05\\
4	9.40273928578491e-05\\
5	9.73131338921825e-05\\
6	9.94886394941543e-05\\
};
\addlegendentry{$N_\mathrm{S} = 32$}

\addplot [color=mycolor3, dotted, line width=2.0pt, mark=square, mark options={solid, mycolor3}]
  table[row sep=crcr]{%
1	9.23552447351784e-05\\
2	9.66298198027635e-05\\
3	0.000102096862854765\\
4	0.000109010799564887\\
5	0.00011684562278114\\
6	0.000121837810358247\\
7	0.000124553664322635\\
};
\addlegendentry{$N_\mathrm{S} = 64$}

\addplot [color=mycolor4, dashdotted, line width=2.0pt, mark=x, mark options={solid, mycolor4}]
  table[row sep=crcr]{%
1	0.000104965329801996\\
2	0.00011389799786835\\
3	0.000126536878725085\\
4	0.000142525515494825\\
5	0.000160952430585978\\
6	0.000173883585167439\\
7	0.000180725719749113\\
8	0.000184115133386408\\
};
\addlegendentry{$N_\mathrm{S} = 128$}

\addplot [color=mycolor5, line width=2.0pt, mark=triangle, mark options={solid, mycolor5}]
  table[row sep=crcr]{%
1	0.000134020224284578\\
2	0.000153465030738225\\
3	0.000181252804375544\\
4	0.000220639327233136\\
5	0.000268738002904876\\
6	0.000305073448328176\\
7	0.000324412126742617\\
8	0.000334907924545081\\
9	0.000340008720154957\\
};
\addlegendentry{$N_\mathrm{S} = 256$}

\end{axis}
\end{tikzpicture}%