### 2024-12-20 assets/simulation/pc_power_sx_txrx16_nd.tex
:construction: Updating structure

% This file was created by matlab2tikz.
%
%The latest updates can be retrieved from
%  http://www.mathworks.com/matlabcentral/fileexchange/22022-matlab2tikz-matlab2tikz
%where you can also make suggestions and rate matlab2tikz.
%
\definecolor{mycolor1}{rgb}{0.00000,0.44706,0.74118}%
\definecolor{mycolor2}{rgb}{0.85098,0.32549,0.09804}%
\definecolor{mycolor3}{rgb}{0.92941,0.69412,0.12549}%
\definecolor{mycolor4}{rgb}{0.49412,0.18431,0.55686}%
\definecolor{mycolor5}{rgb}{0.46667,0.67451,0.18824}%
%
\begin{tikzpicture}

\begin{axis}[%
width=7.607cm,
height=6cm,
at={(0cm,0cm)},
scale only axis,
xmin=1,
xmax=9,
xtick={1,2,3,4,5,6,7,8,9},
xticklabels={{$2^0$},{$2^1$},{$2^2$},{$2^3$},{$2^4$},{$2^5$},{$2^6$},{$2^7$},{$2^8$}},
% xlabel style={font=\color{white!15!black}},
xlabel={RIS Group Size},
ymode=log,
% ymin=0,
ymax=0.000115917330108446,
yminorticks=true,
% ylabel style={font=\color{white!15!black}},
ylabel={Channel Power [W]},
axis background/.style={fill=white},
xmajorgrids,
ymajorgrids,
yminorgrids,
legend style={at={(0.03,0.97)}, anchor=north west, legend cell align=left, align=left, draw=white!15!black}
]
\addplot [color=mycolor1, line width=2.0pt]
  table[row sep=crcr]{%
1	5.62265231393471e-07\\
2	6.30889582242226e-07\\
3	7.11254960263e-07\\
4	7.81326631210281e-07\\
5	8.14448482525861e-07\\
};
\addlegendentry{$N_\mathrm{S} = 16$}

\addplot [color=mycolor2, dashed, line width=2.0pt]
  table[row sep=crcr]{%
1	1.28776059362021e-06\\
2	1.49666859439015e-06\\
3	1.7748873501561e-06\\
4	2.08628203322482e-06\\
5	2.32064182394341e-06\\
6	2.40700699069632e-06\\
};
\addlegendentry{$N_\mathrm{S} = 32$}

\addplot [color=mycolor3, dotted, line width=2.0pt]
  table[row sep=crcr]{%
1	3.15006337524297e-06\\
2	3.87312444339353e-06\\
3	4.86242306926796e-06\\
4	6.08760688271974e-06\\
5	7.25680982863437e-06\\
6	7.92125259488618e-06\\
7	8.12946998504463e-06\\
};
\addlegendentry{$N_\mathrm{S} = 64$}

\addplot [color=mycolor4, dashdotted, line width=2.0pt]
  table[row sep=crcr]{%
1	8.09679926499813e-06\\
2	1.03932623438214e-05\\
3	1.38959563603953e-05\\
4	1.85906493199221e-05\\
5	2.38130628567437e-05\\
6	2.73295770613527e-05\\
7	2.8861300924676e-05\\
8	2.93516301422595e-05\\
};
\addlegendentry{$N_\mathrm{S} = 128$}

\addplot [color=mycolor5, line width=2.0pt]
  table[row sep=crcr]{%
1	2.17722070426804e-05\\
2	2.9526076730669e-05\\
3	4.18291043077588e-05\\
4	5.97087109667717e-05\\
5	8.17036753929845e-05\\
6	9.7962703850956e-05\\
7	0.000105935331911385\\
8	0.000109247922920587\\
9	0.000110397457246139\\
};
\addlegendentry{$N_\mathrm{S} = 256$}

\end{axis}
\end{tikzpicture}%