### 2024-12-30 assets/simulation/pc_power_sx_txrx16_nd.tex
:construction: Add interference channel results

% This file was created by matlab2tikz.
%
%The latest updates can be retrieved from
%  http://www.mathworks.com/matlabcentral/fileexchange/22022-matlab2tikz-matlab2tikz
%where you can also make suggestions and rate matlab2tikz.
%
\definecolor{mycolor1}{rgb}{0.00000,0.44700,0.74100}%
\definecolor{mycolor2}{rgb}{0.00000,0.44706,0.74118}%
\definecolor{mycolor3}{rgb}{0.85000,0.32500,0.09800}%
\definecolor{mycolor4}{rgb}{0.85098,0.32549,0.09804}%
\definecolor{mycolor5}{rgb}{0.92900,0.69400,0.12500}%
\definecolor{mycolor6}{rgb}{0.92941,0.69412,0.12549}%
\definecolor{mycolor7}{rgb}{0.49400,0.18400,0.55600}%
\definecolor{mycolor8}{rgb}{0.49412,0.18431,0.55686}%
\definecolor{mycolor9}{rgb}{0.46600,0.67400,0.18800}%
\definecolor{mycolor10}{rgb}{0.46667,0.67451,0.18824}%
\definecolor{mycolor11}{rgb}{0.30100,0.74500,0.93300}%
\definecolor{mycolor12}{rgb}{0.50196,0.50196,0.50196}%
%
\begin{tikzpicture}

\begin{axis}[%
width=7.607cm,
height=6cm,
at={(0cm,0cm)},
scale only axis,
xmin=1,
xmax=9,
xtick={1,2,3,4,5,6,7,8,9},
xticklabels={{$2^0$},{$2^1$},{$2^2$},{$2^3$},{$2^4$},{$2^5$},{$2^6$},{$2^7$},{$2^8$}},
xlabel style={font=\color{white!15!black}},
xlabel={RIS Group Size},
ymin=0,
ymax=0.000116439758271769,
ylabel style={font=\color{white!15!black}},
ylabel={Channel Power [W]},
axis background/.style={fill=white},
xmajorgrids,
ymajorgrids,
legend style={at={(0.03,0.97)}, anchor=north west, legend cell align=left, align=left, draw=white!15!black}
]
\addplot[only marks, mark=o, mark options={solid}, mark size=2pt, line width=2pt, draw=mycolor2, forget plot] table[row sep=crcr]{%
x	y\\
5	8.18421179472296e-07\\
};
\addplot[only marks, mark=o, mark options={solid}, mark size=2pt, line width=2pt, draw=mycolor4, forget plot] table[row sep=crcr]{%
x	y\\
6	2.47795062804422e-06\\
};
\addplot[only marks, mark=o, mark options={solid}, mark size=2pt, line width=2pt, draw=mycolor6, forget plot] table[row sep=crcr]{%
x	y\\
7	8.20088397806638e-06\\
};
\addplot[only marks, mark=o, mark options={solid}, mark size=2pt, line width=2pt, draw=mycolor8, forget plot] table[row sep=crcr]{%
x	y\\
8	2.94451091720999e-05\\
};
\addplot[only marks, mark=o, mark options={solid}, mark size=2pt, line width=2pt, draw=mycolor10, forget plot] table[row sep=crcr]{%
x	y\\
9	0.000111552731718914\\
};
\addplot[draw=none, only marks, mark=o, mark options={solid}, mark size=2pt, line width=2pt, draw=mycolor12] table[row sep=crcr]{%
x	y\\
0	0\\
};
\addlegendentry{Explicit}

\addplot [color=mycolor2, line width=2.0pt]
  table[row sep=crcr]{%
1	5.66270957115429e-07\\
2	6.32214073230813e-07\\
3	7.11789599032912e-07\\
4	7.83501502399535e-07\\
5	8.1582630412623e-07\\
};
\addlegendentry{$N_\mathrm{S} = 16$}

\addplot [color=mycolor4, dashed, line width=2.0pt]
  table[row sep=crcr]{%
1	1.31853309573688e-06\\
2	1.53662356650387e-06\\
3	1.82774899163227e-06\\
4	2.1412614461017e-06\\
5	2.38070294433103e-06\\
6	2.46743024417076e-06\\
};
\addlegendentry{$N_\mathrm{S} = 32$}

\addplot [color=mycolor6, dotted, line width=2.0pt]
  table[row sep=crcr]{%
1	3.16991712195604e-06\\
2	3.88139289638628e-06\\
3	4.87508772014936e-06\\
4	6.11547438007159e-06\\
5	7.2872459799269e-06\\
6	7.95147687920036e-06\\
7	8.1598841971569e-06\\
};
\addlegendentry{$N_\mathrm{S} = 64$}

\addplot [color=mycolor8, dashdotted, line width=2.0pt]
  table[row sep=crcr]{%
1	8.08840132190028e-06\\
2	1.03513925278208e-05\\
3	1.38370561575957e-05\\
4	1.85667340754753e-05\\
5	2.37528268575109e-05\\
6	2.72839029873604e-05\\
7	2.87929093680483e-05\\
8	2.92781883450922e-05\\
};
\addlegendentry{$N_\mathrm{S} = 128$}

\addplot [color=mycolor10, line width=2.0pt]
  table[row sep=crcr]{%
1	2.18463432534312e-05\\
2	2.96551238699234e-05\\
3	4.20172201243997e-05\\
4	5.99730466582087e-05\\
5	8.20696998678401e-05\\
6	9.8504151463339e-05\\
7	0.00010650814145777\\
8	0.000109789012745988\\
9	0.000110895007877875\\
};
\addlegendentry{$N_\mathrm{S} = 256$}

\end{axis}
\end{tikzpicture}%